xLesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes: In order to observe how audience size varies with differnt features of the person we will use scatterplots

***

### Scatterplots
Notes: Scatterplots let you look at the relationships of 2 variables on one plot

```{r Scatterplots}
getwd()
setwd('/Users/agibbs/code/Udacity/Exploratory Data Analysis/lesson4')

library(ggplot2)
pf = read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
###or
qplot(age, friend_count, data = pf)

```

***

#### What are some things that you notice right away?
Response: Young people tend to have a lot more friends with the exception of a few key lines in later adults (69, and arround 100). Thesea are likely faked ages

***

### ggplot Syntax
Notes: you have to specify the type of plot look back to instructor notes to see the plot types available. Scatterplot is geom_point. 

aes - asthetic wrapper

change the x axis outside the ggplot call with xlim
and add alpha to change the transparency of the data in order to see the overlapping data more clearly

```{r ggplot Syntax}
qplot(x = age, y = friend_count, data = pf)

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() + 
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes: Refers to having too many overlapping poings so that you cannot clearly see the data. This can be addressed in part by changing the transparency (alpha) and adding jitter in order to disperse the points a bit

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90)

```

#### What do you notice in the plot?
Response:
At young ages the distibution approximates a normal with a long upper tail along with a signifcant spike at ~69. Most of young users actually have less than 1000 friends as opposed to what originally seemed to be much higher
***

### Coord_trans()
Notes:read coord_trans info

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

#### What do you notice?
There are very few points above 2000
***

### Alpha and Jitter
Notes: You have to be careful to only jitter ages. If you add noise to zero friend counts the sqrt is imaginary

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

***
### Alpha and Jitter
Explore the relationship between friends initiated vs age

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/10, position = position_jitter(h=0)) + 
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes: Transform the axis to be as a percentage of their total friends

***

### Conditional Means

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

#filter()
#group_by()
#mutate()
#arrange()

age_groups = group_by(pf, age)

pf.fc_by_age = summarise(age_groups, friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n())

head(pf.fc_by_age)
pf.fc_by_age = arrange(pf.fc_by_age, age)

head(pf.fc_by_age)


###alternate code:
pf.fc_by_age = pf %.%
  group_by(age) %.%
  summarise(friend_count_mean = mean(friend_count), 
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age)
head(pf.fc_by_age, 20)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line() + 
  xlab('Age') + ylab('Mean Friend Count')

ggplot(aes(x = age, y = friend_count_median), data = pf.fc_by_age) + 
  geom_line() + 
  xlab('Age') + ylab('Mean Friend Count')
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0), color = 'orange') + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt') + 
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .1, 
            linetype = 2, color = 'blue')  + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .9, 
            linetype = 2, color = 'blue')   + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .5, 
            color = 'blue')


ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0), color = 'orange') + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0,1000)) + 
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .1, 
            linetype = 2, color = 'blue')  + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .9, 
            linetype = 2, color = 'blue')   + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .5, 
            color = 'blue')

#add coord_cartesian to zoom in
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes: Correlation coefficient

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count))

with(subset(pf, age<=70), cor.test(age, friend_count, method = 'spearman'))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha = 1/5)

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha = 1/20) + 
  coord_cartesian(xlim = c(0,5000), ylim = c(0,15000))

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(0,quantile(pf$www_likes_received, 0.95)) +
  ylim(0,quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell

```

Create your plot!

```{r Temp vs Month}
names(Mitchell)
ggplot(aes(Month, Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)

ggplot(aes(Month, Temp), data = Mitchell) +
  geom_point() +
  scale_x_discrete(breaks = seq(0,203, 12))

```

***

### A New Perspective

What do you notice?
Response: For each year there is a similar pattern for temperature which starts out low and rises then comes back down (presumably, beginning in Jan and ending in Dec)

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + 
  geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
names(pf)

pf$age_with_months = pf$age + (12-pf$dob_month)/12
```

Programming Assignment
```{r Programming Assignment}

pf.fc_by_age_months = pf %.%
  group_by(age_with_months) %.%
  summarise(friend_count_mean = mean(friend_count), 
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age_with_months)


```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(age_with_months, friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71))+
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 = ggplot(aes(age, friend_count_mean), 
            data = subset(pf.fc_by_age, age < 71)) + 
  geom_line() + 
  geom_smooth()

p2 = ggplot(aes(age_with_months, friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71))+
  geom_line() +
  geom_smooth()

p3 = ggplot(aes(round(age/5)*5, friend_count), 
            data = (subset(pf, age < 71))) + 
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

