Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: She tried to add age to perceived views and did not find a significant correlation

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
getwd()
setwd('/Users/agibbs/code/Udacity/Exploratory Data Analysis//lesson5')

pf=read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape=4)

ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library(dplyr)
#chain functions together %.%
pf.fc_by_age_gender = pf %.%
  filter(!is.na(gender)) %.%
  group_by(age, gender) %.%
  summarise(mean_friend_count = mean(friend_count), 
            median_friend_count = median(friend_count),
            n = n()) %.%
  ungroup() %.%
  arrange(age)

head(pf.fc_by_age_gender)

ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:
How many times more friends does the average female friends have then male users
***

### Wide and Long Format
Notes:
long format means many rows current
reshape to have wide format
age, male, female are the columns
***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide = dcast(pf.fc_by_age_gender, 
                                 age ~ gender,
                                 value.var = 'median_friend_count') 
#d for data frame; acast for array or matrix
head(pf.fc_by_age_gender.wide)

#play with melt function to go from wide to long format
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.wide$ratioFM = 
  pf.fc_by_age_gender.wide$female/pf.fc_by_age_gender.wide$male

ggplot(aes(x = age, y = ratioFM), data = pf.fc_by_age_gender.wide) + 
  xlab('Age') + ylab('Female/Male Median Friend Count') +
  geom_line() + 
  geom_hline(yintercept=1, alpha = 0.3, color = 'red', linetype = 2)
```

***

### Third Quantitative Variable
Notes: Exploring more variables using a 2 dimensional splay like a scatterplot

```{r Third Quantitative Variable}
summary(pf$tenure)
pf$year_joined = floor(2014-pf$tenure/365)
head(pf$year_joined)
summary(pf$year_joined)
```

***

### Cut a Variable
Notes: make discrete variables to create bins of users

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket = cut(pf$year_joined, c(2004,2009, 2011, 2012, 2014))
pf$year_joined.bucket
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 3)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf$friend_rate = pf$friend_count/pf$tenure
summary(subset(pf$friend_rate, pf$tenure>=1))

with(subset(pf, tenure>=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure), 
       data = subset(pf, tenure>=1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)

ggplot(aes(x = 30*round(tenure/30), y = friendships_initiated/tenure), 
       data = subset(pf, tenure>=1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket)) 
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo = read.csv('yogurt.csv')
names(yo)
yo$id = factor(yo$id)
str(yo)

ggplot(aes(price), data = yo) + geom_histogram() + theme_minimal()
summary(yo$price)
length(unique(yo$price))
table(yo$price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo = transform(yo, all.purchases = strawberry + blueberry +
                            pina.colada + plain + mixed.berry)

ggplot(aes(all.purchases), data = yo) + geom_histogram(bin = 1) + 
  theme_minimal()
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(time, price), data = yo) + geom_jitter(alpha = 1/5) + 
  theme_minimal()
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids = sample(levels(yo$id), 16)
sample.ids

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)


set.seed(5)
sample.ids = sample(levels(yo$id), 16)
sample.ids

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***
### Scatterplot Matrix
```{r}
#install.packages('GGally')
library(GGally)
theme_set(theme_minimal())

#set the seed for reproducible results
set.seed(1836)
pf_subset = pf[, c(2,7,8,9)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

