Problem Set 3
========================================================
  
***
  
### Load ggplot library and diamonds data set

```{r}
setwd('/Users/agibbs/code/Udacity/Exploratory Data Analysis/ProblemSet3/')
library(ggplot2)
data(diamonds)
names(diamonds)
str(diamonds)
diamonds$color
```


***
  
### Histogram of price of all the diamonds
```{r Pseudo-Facebook User Data}
qplot(price, data = diamonds, binwidth = 100)
```


### Summary of diamond price
```{r}
summary(diamonds$price)

sum(diam500$price < 500)

sum(diamonds$price < 250)

sum(diamonds$price >= 15000)
```

### Histogram of price of lowest peak of the diamonds
```{r Pseudo-Facebook User Data}
qplot(price, data = diamonds, binwidth = 10) + 
  scale_x_continuous(limits = c(300,1500), breaks = seq(0,6000, 50))
```

### Break out the histogram of diamond prices by cut
### end up with 5 histograms in separate panels on the plot
```{r}
qplot(price, data = diamonds, binwidth = 100, 
      xlab = 'Diamond Price ($)', ylab = 'Number of Diamonds') +
  scale_x_continuous(breaks = seq(0,150000, 5000)) + 
  facet_wrap(~cut, ncol = 1)

by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$cut, max)
by(diamonds$price, diamonds$cut, min)
by(diamonds$price, diamonds$cut, median)

tapply(diamonds$price, diamonds$cut, summary)
```

```{r}
qplot(price, data = diamonds, binwidth = 200, 
      xlab = 'Diamond Price ($)', ylab = 'Number of Diamonds') + 
  scale_x_continuous(limits = c(0,20000), 
                     breaks = seq(0,150000, 5000)) +   
  facet_wrap(~cut, ncol = 1, scales = 'free_y')
```

### Diamond Price/Carat
```{r}
qplot(price/carat, data = diamonds, binwidth = 0.02,
      color = I('black'), fill = I('#5760AB'),
      xlab = 'Diamond Price/Carat', ylab = 'Number of Diamonds') + 
  scale_x_log10() +
  facet_wrap(~cut, ncol = 1, scales = 'free_y')
```


### Investigate the price of diamonds using box plots, numerical summaries and the following categorical variables: cut, clarity, and color
#by Cut, clarity and color
```{r}
qplot(x = cut, y = price, 
      data = diamonds, 
      geom = 'boxplot')+
  coord_cartesian()

qplot(x = cut, y = price, 
      data = diamonds, 
      geom = 'boxplot')+
  coord_cartesian(ylim = c(0,7000))


tapply(diamonds$price, diamonds$cut, summary)

qplot(x = clarity, y = price, 
      data = diamonds, 
      geom = 'boxplot') + 
  coord_cartesian()

qplot(x = clarity, y = price, 
      data = diamonds, 
      geom = 'boxplot')+
  coord_cartesian(ylim = c(0,7000))

tapply(diamonds$price, diamonds$clarity, summary)

qplot(x = color, y = price, 
      data = diamonds, 
      geom = 'boxplot') +
  coord_cartesian()

qplot(x = color, y = price, 
      data = diamonds, 
      geom = 'boxplot')+
  coord_cartesian(ylim = c(0,8000))

by(diamonds$price, diamonds$color, summary)

IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)

qplot(x = color, y = price/carat, 
      data = diamonds, 
      geom = 'boxplot') +
  coord_cartesian()

qplot(x = color, y = price/carat, 
      data = diamonds, 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0,6000))

qplot(x = color, y = price/carat, 
      data = diamonds, 
      geom = 'boxplot') +
  coord_cartesian() +
  facet_wrap(~clarity)

qplot(x = color, y = price/carat, 
      data = diamonds, 
      geom = 'boxplot') +
  coord_cartesian() +
  facet_wrap(~cut)
```


###Investiage carat using frequency polygons
```{r}
qplot(x = carat, 
      data = diamonds, 
      xlab = 'Carat',
      ylab = 'Number of Diamonds',
      binwidth = 0.01, geom = 'freqpoly') +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0, 1, 0.1)) 

qplot(x = carat, 
      data = diamonds, 
      xlab = 'Carat',
      ylab = 'Number of Diamonds',
      binwidth = 0.01, geom = 'freqpoly', color = color) +
  scale_x_continuous(limits = c(0,1)) 
```

###Investigate breast cancer data from Gapminder
```{r}
BC = read.csv('BreastCancerMortaility.csv', na.strings='NA', header = T, row.names=1, blank.lines.skip=T)

qplot(x = X2002, data = BC)

#read in data on health care spending
HSpend = read.csv('HealthSpending.csv', na.strings='NA', header=T, row.names=1, blank.lines.skip=T)

#setup column to merge on
BC$country = rownames(BC)
HSpend$country = rownames(HSpend)

#merge data sets by country
BCandHS = merge(BC, HSpend, by='country')

#visualize each dataset for 2002
qplot(x = X2002.x, data = BCandHS)
qplot(x = X2002.y, data = BCandHS)

#look at summary of health care spending data set to decide how to split
summary(BCandHS$X2002.y)

#split healthcare spending for 2002 and create new column with very low, low, high, and very high spending
BCandHS$HS = NA
BCandHS$HS = ifelse(BCandHS$X2002.y <= 24.21, 'very low', NA)
BCandHS$HS = ifelse(BCandHS$X2002.y > 24.21 & BCandHS$X2002.y <= 105.00, 'low', BCandHS$HS)
BCandHS$HS = ifelse(BCandHS$X2002.y > 105.00 & BCandHS$X2002.y <= 376.10, 'high', BCandHS$HS)
BCandHS$HS = ifelse(BCandHS$X2002.y > 376.10, 'very high', BCandHS$HS)

#set order of levels for health care spending
hsLevels = c("very low", "low", "high", "very high")
BCandHS$HS = factor(BCandHS$HS, levels = hsLevels)

#create histograms of breast cancer mortaility based on levels of health care spending
qplot(x = X2002.x, data = subset(BCandHS, !is.na(X2002.y)), 
      binwidth = 1) +
  scale_x_continuous() + 
  facet_wrap(~HS)

#create frequency polygons 
qplot(x = X2002.x, y = ..count../sum(..count..), 
      data = subset(BCandHS, !is.na(X2002.y)), 
      xlab = 'Breast Cancer Mortality',
      ylab = 'Proportion of countries',
      binwidth = 4, geom = 'freqpoly', color = HS) +
  scale_x_continuous() 

#create boxplots
qplot(x = HS, y = X2002.x, 
      data = subset(BCandHS, !is.na(X2002.y)),
      xlab = 'Health Spending',
      ylab = 'Breast Cancer Mortaility',
      geom = 'boxplot')
```



###aquire data from my facebook account on friend's birthdays and investigate:
#how many people share my bday?
#which month contains the most birthdays?
#how many birthdays are in each month?
#which day of the year has the most birthdays?
#what day of the month has the most birthdays?
#what days/months of the year have the least bdays?

```{r}
bdays = read.csv('Bdays.csv', sep = '\t')

library('lubridate')

bdays['Date'] = mdy(bdays[,2])
bdays['Month'] = month(bdays[,2])
bdays['Day'] = mday(bdays[,2])
bdays['DayY'] = yday(bdays[,2])

##find the people who share my birthday
MyBdayDofY = yday(mdy('03/22/2014'))
ShareBday = subset(bdays, bdays['DayY'] == MyBdayDofY)
ShareBday['Name']
#No one shares my birthday!
BdayM1 = subset(bdays, bdays['DayY'] == MyBdayDofY-1)
BdayM1['Name']
#2 people (Paul and Zsuzsanna) have a birthday the previous day

#determine which month has the most birthdays
qplot(x = Month, data=bdays) + 
  scale_x_discrete(breaks = seq(1, 12, 1)) + 
  xlab('Month') + 
  ylab('Frequency')
table(bdays$Month)
#March has the most birthdays followed by December, November, September, and February
# May and August tie for the least number of birthdays

# Month:      1  2  3  4  5  6  7  8  9 10 11 12 
# # of bdays: 18 22 27 18 16 23 17 16 23 19 24 26 

#determine which day of the year has the most birthdays
qplot(x = DayY, data=bdays, binwidth = 1) + 
  scale_x_discrete(breaks = seq(100, 150, 5)) + 
  xlab('Day') + 
  ylab('Frequency')
table(bdays$DayY)
#there is one day with 3 total birthdays and several with 3
TalyDays = table(bdays$DayY)
CommonDays = subset(TalyDays, TalyDays>2)
CommonDays
#day 348 - 12/14 has the most bdays (4)

qplot(x = Day, data=bdays, binwidth = 1) + 
  scale_x_discrete(breaks = seq(1,31,1)) + 
  xlab('Day of the Month') + 
  ylab('Frequency')
table(bdays$Day)
#the most birthdays fall in the middle of the month on the 14th the 31st, 24th, 3rd, and 6th tie for the least. This is a little surprising since only half of the months have a 31st day

qplot(x = DayY, data=bdays, binwidth = 1) + 
  scale_x_continuous(limits = c(114, 130), breaks = seq(100, 150, 5)) + 
  xlab('Day') + 
  ylab('Frequency')
#I do not have 365 friends and there are many days which do not have birthdays. The largest gap with no birthdays is from day 115 to 129
```


***