Exploratory Data Analysis: Problem Set 4
========================================================


```{r}
setwd('/Users/agibbs/code/Udacity/Exploratory Data Analysis/ProblemSet4/')
library(ggplot2)
data(diamonds)
names(diamonds)
```

###Create a scatterplot of diamond price versus x

```{r}
ggplot(aes(x, price), data = diamonds) + 
  geom_point()

cor.test(diamonds$x, diamonds$price)
cor.test(diamonds$y, diamonds$price)
cor.test(diamonds$z, diamonds$price)
```

#create a simple scatter plot of price vs depth

```{r}
ggplot(aes(depth, price), data = diamonds) +
  geom_point(alpha = 1/100) + 
  scale_x_continuous(breaks = seq(43, 79, 2))

range(diamonds$depth)

cor.test(diamonds$depth, diamonds$price)
```

#create a simple scatter plot of price vs depth

```{r}
ggplot(aes(carat, price), data = diamonds) +
  geom_point() +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```

#Create a scatterplot of price vs volume (x*y*z)
```{r}
diamonds$volume = diamonds$x*diamonds$y*diamonds$z

ggplot(aes(volume, price), data = diamonds) + 
  geom_point()

subset(diamonds, diamonds$volume>800)

library(plyr); library(dplyr)
count(diamonds$volume ==0)

detach("package:plyr", unload=TRUE)

Diamonds_less_vol_outliers = subset(diamonds, volume <= 800 & volume >0)

cor.test(Diamonds_less_vol_outliers$volume, Diamonds_less_vol_outliers$price)

ggplot(aes(volume, price), data = Diamonds_less_vol_outliers) + 
  geom_point(alpha = 1/30) +
  geom_smooth(method = 'lm', color = 'red')

ggplot(aes(volume, price), data = Diamonds_less_vol_outliers) + 
  geom_point(alpha = 1/30) +
  geom_smooth(method = 'lm', color = 'red') +
  geom_smooth(formula = y ~x^3, color = 'blue') +
  xlim(0, 300) + 
  ylim(0,11000)

ggplot(aes(volume, price), data = Diamonds_less_vol_outliers) + 
  geom_point(alpha = 1/20) +
  scale_y_sqrt() +
  geom_smooth(method = 'lm', color = 'red')
```

###Examine diamonds by clarity more closely
```{r}
library(dplyr)
diamondsByClarity = diamonds %.%
  group_by(clarity) %.%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n=n()) %.%
  arrange(clarity)

```

###Create 2 bar plots on one output image using grid.arrange() showing mean prices by clarity and color
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

cl = ggplot(diamonds_mp_by_clarity, aes(clarity, mean_price)) + 
  geom_bar(stat='identity')
co = ggplot(diamonds_mp_by_color, aes(color, mean_price)) + 
  geom_bar(stat='identity')

grid.arrange(cl, co, ncol =1)

diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

cu = ggplot(diamonds_mp_by_cut, aes(cut, mean_price)) + 
  geom_bar(stat='identity') 
grid.arrange(cl, co, cu, ncol=1)
```

###Gapminder Data exploring Breast Cancer Mortality, Incidence and health care spending

```{r}
library(gridExtra)

#read in data on Breast Cancer Mortality
BCm = read.csv('BreastCancerMortaility.csv', na.strings='NA', header = T, row.names=1, blank.lines.skip=T)

#read in data on Breast Cancer Incidence
BCi = read.csv('BreastCancerIncidence.csv', na.strings='NA', header = T, row.names = 1, blank.lines.skip = T)

#read in data on health care spending
HSpend = read.csv('HealthSpending.csv', na.strings='NA', header=T, row.names=1, blank.lines.skip=T)

#setup column to merge on
BCm$country = rownames(BCm)
BCi$country = rownames(BCi)
HSpend$country = rownames(HSpend)

#merge data sets by country for 2002
BCmandHS = merge(BCm[,c('country','X2002')], HSpend[,c('country', 'X2002')], by='country')

BCandHS = merge(BCi[,c('country','X2002')], BCmandHS, by='country')

colnames(BCandHS) = c('country','Cancer_Incidence', 'Cancer_Mortality', 'Health_Spending')

#plot Breast Cancer incidence and mortaility vs Health care spending
inc = ggplot(BCandHS, aes(Health_Spending, Cancer_Incidence)) + 
  geom_point() + 
  xlab('Health Care Spending') +
  ylab('Breast Cancer Incidence')
mort = ggplot(BCandHS, aes(Health_Spending, Cancer_Mortality)) + 
  geom_point() + 
  xlab('Health Care Spending') +
  ylab('Breast Cancer Mortality')
grid.arrange(inc, mort)

#test correlation
cor.test(BCandHS$Health_Spending, BCandHS$Cancer_Incidence)
cor.test(BCandHS$Health_Spending, BCandHS$Cancer_Mortality)

#log breast cancer incidence data and fit smoothed curve
ggplot(BCandHS, aes(Health_Spending, Cancer_Incidence)) + 
  geom_point() + 
  scale_x_log10() +
  geom_smooth(method = 'loess', span = 0.5, color = 'blue') + 
  ylab('Breast Cancer Incidence') + 
  xlab('Healthcare Spending')

#fit linear model to log10 data
inc10 = ggplot(BCandHS, aes(Health_Spending, Cancer_Incidence)) + 
  geom_point() + 
  scale_x_log10() +
  geom_smooth(method = 'lm', color = 'blue') + 
  ylab('Breast Cancer Incidence') + 
  xlab('Healthcare Spending')

mort10 = ggplot(BCandHS, aes(Health_Spending, Cancer_Mortality)) + 
  geom_point() +
  scale_x_log10() + 
  geom_smooth(method = 'lm', color = 'blue') + 
  ylab('Breast Cancer Mortality') + 
  xlab('Healthcare Spending')

grid.arrange(inc10, mort10)

#break down health care spending into levels of spending
BCandHS$HS = NA
BCandHS$HS = ifelse(BCandHS$Health_Spending <= 24.21, 'very low', NA)
BCandHS$HS = ifelse(BCandHS$Health_Spending > 24.21 & BCandHS$Health_Spending <= 105.00, 'low', BCandHS$HS)
BCandHS$HS = ifelse(BCandHS$Health_Spending > 105.00 & BCandHS$Health_Spending <= 376.10, 'medium', BCandHS$HS)
BCandHS$HS = ifelse(BCandHS$Health_Spending > 376.10 & BCandHS$Health_Spending <= 1000, 'high', BCandHS$HS)
BCandHS$HS = ifelse(BCandHS$Health_Spending > 1000, 'high', BCandHS$HS)

hsLevels = c("very low", "low", "medium","high", "very high")
BCandHS$HS = factor(BCandHS$HS, levels = hsLevels)

BCbyHS = BCandHS %.%
  group_by(HS) %.%
  summarise(BCi_mean = mean(Cancer_Incidence), 
            BCm_mean = mean(Cancer_Mortality), 
            BCi_median = median(Cancer_Incidence), 
            BCm_median = median(Cancer_Mortality), 
            n = n()) %.%
  arrange(HS)

p1 = ggplot(BCbyHS, aes(HS, BCi_mean)) +
  geom_bar(stat='identity') + 
  xlab('Breast Cancer Incidence') + 
  ylab('Healthcare Spending')

p2 = ggplot(BCbyHS, aes(HS, BCm_mean)) +
  geom_bar(stat='identity') + 
  xlab('Breast Cancer Mortality') + 
  ylab('Healthcare Spending')

grid.arrange(p1, p2)
```

